---
title: "Timeseries workflow"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
vignette: >
  %\VignetteIndexEntry{Timeseries workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This vignette describes a general workflow for querying and viewing data from the Aquarius database using the `deltawqAQ` package. Steps include identifying the appropriate location and parameter codes using the `aq_all_locations` dataset and the `aq_get_location_parameters` function, then querying a timeseries using the `aq_get_ts` function, and making a basic plot. For working with multiple parameters or locations, see the `retrieve_multi_ts` vignette. 

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Start by loading packages. 

```{r setup, warning = FALSE, message = FALSE}
library(deltawqAQ)
library(ggplot2)
library(lubridate)
library(dplyr)
library(DT)
library(leaflet)
```

To work with Aquarius data, you may want to start with exploring available locations and/or available parameters.

## Start with location

You may start by figuring out the location code(s) of interest. Functions will take `cdec_code`, `location_id`, or `aquarius_id`, with most users tending to be most familiar with `cdec_code`. 

Use the `aq_all_locations` data object to view all the available locations. This object gets generated in the package, and locations should be stable once all locations become telemetered.

```{r}
DT::datatable(deltawqAQ::aq_all_locations, 
              rownames = FALSE,
              options = list(scrollX = TRUE,
                             order = list(list(1, 'asc')),
                             autoWidth = TRUE))
```

View locations on a map. 

```{r}
leaflet() %>%
  addTiles() %>%
  addMarkers(lng = deltawqAQ::aq_all_locations$longitude, 
             lat = deltawqAQ::aq_all_locations$latitude, 
             label = deltawqAQ::aq_all_locations$cdec_code)
```

### Look at available parameters
Once you have picked your location, take a look at the parameters you would like to query. The `aq_get_location_parameters` function will connect to the Aquarius database based on your credentials, then retrieve all parameters associated with the location you have selected. 

We have selected the `SJW` location for this example. You will need `parameter_name` for your time series query, so identify the correct value for the next step.

```{r}
param_list <- aq_get_location_parameters(cdec_code = "SJW")
DT::datatable(param_list, 
              rownames = FALSE,
              options = list(scrollX = TRUE, autoWidth = TRUE))
```

## Start with parameter

Instead of starting with location, you may also start by figuring out the parameter code(s) of interest. 

```{r}
DT::datatable(deltawqAQ::aq_all_parameters, 
              rownames= FALSE,
              options = list(scrollX = TRUE,
                             order = list(list(1, 'asc'))))
```

### Look at available locations

From there you can look at which locations have data for your parameter of interest

```{r}
wt_locations <- aq_get_parameter_locations("Water Temp")
DT::datatable(wt_locations, 
              rownames= FALSE,
              options = list(scrollX = TRUE,
                             autoWidth = TRUE))
```

## Query data

Now you can run your query for your location and parameter of interest. You can enter a datetime or a date for the `query_from` and `query_to` fields. 

```{r}
ts <- aq_get_ts(cdec_code = "SJW", 
                parameter = "Water Temp", 
                query_from = "2025-12-01 00:00:00", 
                query_to = "2026-02-01 00:00:00")
```

```{r}
DT::datatable(ts, 
              rownames = FALSE, 
              options = list(scrollX = TRUE, autoWidth = TRUE))
```

## Summarize data
You may want to condense the data by creating a dataset with daily averages. 

```{r, warning = FALSE, message = FALSE}
ts_daily <- ts |> 
  mutate(date = lubridate::date(datetime)) |>
  group_by(date, cdec_code) |>
  summarize(wt_daily = round(mean(value),2)) |>
  ungroup()
```

```{r}
DT::datatable(ts_daily, 
              rownames = FALSE,
              options = list(scrollX = TRUE))
```

## Visualize data
Now you can make a simple plot for daily average values for the time series.

```{r, fig.height = 4, fig.width = 6, fig.title = ts_daily$aq_location_name[1], fig.alt = "Plot of daily average water temperature at SJW"}

ggplot(ts_daily) + 
  geom_point(aes(x = date, y = wt_daily)) +
  geom_line(aes(x = date, y = wt_daily)) +
  labs(x = "Date", y = "Water Temp (°C)") +
  theme_bw()+
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 11))
```

### Write data

You can then write data to your local drive. Specify the filepath of interest and customize the filename if desired. 

```{r, eval = FALSE}
write.csv(ts_daily, 
          here::here(paste0("ts_SJW_wt.csv")))
```
