---
title: "Timeseries workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Timeseries workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This vignette describes a general workflow for querying and viewing data from the Aquarius database using the `deltawqAQ` package. Steps include identifying the appropriate location and parameter codes using the `aq_all_locations` dataset and the `aq_get_station_parameters` function, then querying a timeseries using the `aq_get_ts` function, and making a basic plot. For working with multiple parameters or stations, see the `retrieve_multi_ts` vignette. 

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Start by loading packages. 

```{r setup}
library(deltawqAQ)
library(ggplot2)
library(lubridate)
library(dplyr)
library(knitr)
```

## Locations

To work with Aquarius data, start by figuring out the location code(s) of interest. Functions will take `cdec_code`, `location_id`, or `aquarius_id`, with most users tending to be most familiar with `cdec_code`. The `aq_all_locations` data object gets generated in the package, and locations should be stable once all locations become telemetered.

```{r, echo = FALSE}
kable(head(deltawqAQ::aq_all_locations))
```

## Parameters
Once you have picked your station, take a look at the parameters you would like to query. We have selected `SJW` for this example. You will need `parameter_name` for your timeseries query.

```{r}
param_list <- aq_get_station_parameters(cdec_code = "SJW")
```

```{r echo = FALSE}
kable(head(param_list, n = 15))
```

## Querying

Now you can run your query for your station and parameter of interest. You can enter a datetime or a date for the `query_from` and `query_to` fields. 

```{r}
ts <- aq_get_ts(cdec_code = "SJW", 
                parameter = "Water Temp", 
                query_from = "2025-12-01 00:00:00", 
                query_to = "2026-02-01 00:00:00")
```

```{r, echo = FALSE}
kable(head(ts))
```

## Modify data

You may want to condense the data by getting daily averages. 

```{r}
ts_daily <- ts |> 
  mutate(date = lubridate::date(datetime)) |>
  group_by(date, cdec_code) |>
  summarize(wt_daily = mean(value)) |>
  ungroup()
```

```{r, echo = FALSE}
kable(head(ts_daily))
```

## View data
Now you can make a simple plot for daily average values for the timeseries.

```{r, fig.height = 4, fig.width = 6}
ggplot(ts_daily) + 
  geom_point(aes(x = date, y = wt_daily)) +
  geom_line(aes(x = date, y = wt_daily)) +
  labs(x = "Date", y = "Water Temp (°C)", title = ts_daily$aq_location_name[1] ) +
  theme_bw()+
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 11))
```


