---
title: "Retrieve multiple parameters and stations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Retrieve multiple parameters and stations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette describes a general workflow for querying and viewing multi-parameter and multi-station data from the Aquarius database using the `deltawqAQ` package. Steps include identifying the appropriate location and parameter codes using the `aq_all_locations` dataset and the `aq_get_station_parameters` function, then querying timeseries using the `aq_get_ts_multi_param` and `aq_get_ts_multi_station` functions. Code for making basic plot and writing the data are also included. 
```{r setup}
library(deltawqAQ)
library(ggplot2)
library(lubridate)
library(dplyr)
library(knitr)
```


## Select Station

To work with Aquarius data, start by figuring out the location code(s) of interest. Functions will take `cdec_code`, `location_id`, or `aquarius_id`, with most users tending to be most familiar with `cdec_code`. The `aq_all_locations` data object gets generated in the package, and locations should be stable once all locations become telemetered.

```{r}
kable(head(deltawqAQ::aq_all_locations))
```

## Multiple parameters for one station
Once you have picked your station, take a look at the parameters you would like to query. We have selected `SJW` for this example. You will need `parameter_name` for your timeseries query.

```{r}
station = "SJW"
```

### Select parameter names
Now use `aq_get_station_parameters` to figure out the parameter names for your station.

```{r, echo = FALSE}
kable(head(aq_get_station_parameters(cdec_code = station), n = 15))
```

### Query data 
Now you can use the `aq_get_ts_multi_param` function to run your query for your station and parameters of interest. You can enter a datetime or a date for the `query_from` and `query_to` fields. 

#### Select parameters and dates
```{r}
params = c("Turbidity, Form Neph", "Water Temp", "Sp Cond", "CHL RFU", "pH")
start_date = "2025-12-01"
end_date = lubridate::today()
```

#### Query and look at data
```{r}
multi_param_ts <- aq_get_ts_multi_param(
  cdec_code = station, 
  parameter = params, 
  query_from = start_date, 
  query_to = end_date
  )

print(paste("Parameters with data:", paste(unique(multi_param_ts$parameter_name), collapse = ", ")))
```

```{r, echo = FALSE}
kable(head(multi_param_ts))
```


### Summarize by daily values
You may want to condense the data by getting daily averages.

```{r}
multi_param_daily <- multi_param_ts %>%
  mutate(date = lubridate::date(datetime)) %>%
  group_by(date, cdec_code, location_id, parameter_name) %>%
  summarize(daily_avg = mean(value)) %>%
  ungroup()
```

```{r, echo = FALSE}
kable(head(multi_param_daily))
```


### Visualize time series for each parameter

If you'd like to visualize your data, here are some examples of plots you can make with your data. 

#### Daily Averages

```{r fig.height = 7, fig.width = 6}
ggplot(multi_param_daily) +
  geom_point(aes(x=date, y=daily_avg, color = parameter_name)) +
  geom_line(aes(x=date, y=daily_avg, color = parameter_name)) +
  facet_wrap(~parameter_name, nrow = nrow(multi_param_daily), scales = "free_y") +
  labs(x = "Date", y = "Daily Average Value", title = paste("Water Quality for", station))+
  theme_bw() +
  theme(strip.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 11))
```

#### 15 minute data

```{r, fig.height = 7, fig.width = 6}
ggplot(multi_param_ts) +
  geom_line(aes(datetime, value, color = parameter_name)) +
  facet_wrap(~parameter_name, nrow = nrow(multi_param_ts), scales = "free_y") +
  labs(x = "Datetime", y = "Value", title = paste("Water Quality for", station))+
  theme_bw() +
  theme(strip.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 11))
```

### Write data

You can then write data to your local drive. Specify the filepath of interest and customize the filename if desired. This example writes the averaged data. 

```{r}
param_label <- paste(unique(multi_param_ts$parameter_name), collapse = "-")

write.csv(multi_param_daily, 
          here::here(paste0("ts_", station, "_",param_label,"_",  start_date, "-", end_date)))
```

## Multiple stations for one parameter

Once you have picked your stations, take a look at the parameter you would like to query. We have selected several stations for this example. You will need `parameter_name` for your timeseries query.

```{r}
stations <- c("GSS", "SMR", "OSJ", "HOL", "OBI", "OH4", "MDM")
```

### Select parameter names

Now use `aq_get_station_parameters` to figure out the parameter names for your station.

```{r, echo = FALSE}
kable(head(aq_get_station_parameters(cdec_code = stations), n = 15))
```

### Query data 
Now you can use the `aq_get_ts_multi_station` function to run your query for your parameter and stations of interest. You can enter a datetime or a date for the `query_from` and `query_to` fields. 

#### Select query parameter and dates
```{r}
param = "Turbidity, Form Neph"
startdate = "2025-12-01"
enddate = lubridate::today()
```

#### Query and look at data
```{r}
multi_station_ts <- aq_get_ts_multi_station(cdec_code = stations, 
                                         parameter = param, 
                                         query_from = startdate, 
                                         query_to = enddate)

print(paste("Stations with data:", paste(unique(multi_station_ts$cdec_code), collapse = ", ")))
```

```{r, echo = FALSE}
kable(head(multi_station_ts))
```

### Summarize by daily values
You may want to condense the data by getting daily averages.

```{r}
multi_station_daily <- multi_station_ts |> 
  mutate(date = lubridate::date(datetime)) |> 
  group_by(date, cdec_code, location_id, parameter_name) |> 
  summarize(daily_avg = mean(value)) |> 
  ungroup()
```

```{r, echo = FALSE}
kable(head(multi_station_daily))
```


### Visualize time series for each station

If you'd like to visualize your data, here are some examples of plots you can make with your data. 

#### Daily Averages

```{r, fig.height = 7, fig.width = 6}
ggplot(multi_station_daily) +
  geom_point(aes(x=date, y=daily_avg, color = cdec_code)) +
  geom_line(aes(x=date, y=daily_avg, color = cdec_code)) +
  facet_wrap(~cdec_code, nrow = nrow(multi_station_daily), scales = "free_y") +
  labs(x = "Date", y = "Daily Average Value", title = paste("Water Quality for", param))+
  theme_bw() +
  theme(strip.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 11))
```

#### 15 minute data

Include a threshold line if useful.

```{r, fig.height = 7, fig.width = 6}
ggplot(multi_station_ts) +
  geom_line(aes(datetime, value, color = cdec_code)) +
  geom_hline(yintercept = 12, linetype = "dashed") + 
  facet_wrap(~cdec_code, nrow = nrow(multi_station_ts), scales = "free_y") +
  labs(x = "Datetime", y = param, color = "station")+
  theme_bw() +
  theme(strip.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.position = "none",
        axis.text = element_text(size = 11))
```

### Write data

You can then write data to your local drive. Specify the filepath of interest and customize the filename if desired. 

```{r}
sta_label <- paste(unique(multi_station_ts$cdec_code), collapse = "-")
write.csv(multi_station_daily, 
          here::here(paste0("ts_", param, "_",sta_label,"_",  start_date, "-", end_date)))
```

## Combine parameter and station

