---
title: "Retrieve timeseries"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
vignette: >
  %\VignetteIndexEntry{Retrieve multiple parameters and locations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette describes a general workflow for querying and viewing multi-parameter and multi-location data from the Aquarius database using the `deltawqAQ` package. Steps include identifying the appropriate location and parameter codes using the `aq_all_locations` dataset and the `aq_get_location_parameters` function, then querying timeseries using the `aq_get_ts` function, which can query a single parameter and location, a single parameter and multiple locations, or a single location with multiple parameters. Code for making basic plot and writing the data are also included. 

```{r setup, warning = FALSE, message = FALSE}
library(deltawqAQ)
library(ggplot2)
library(lubridate)
library(dplyr)
library(DT)
```

## Single location or parameters

To work with Aquarius data, you may want to start with exploring available locations and/or available parameters.

### Start with location

You may start by figuring out the location code(s) of interest. Functions will take `cdec_code`, `location_id`, or `aquarius_id`, with most users tending to be most familiar with `cdec_code`. 

Use the `aq_all_locations` data object to view all the available locations. This object gets generated in the package, and locations should be stable once all locations become telemetered.

```{r}
DT::datatable(deltawqAQ::aq_all_locations, 
              rownames = FALSE,
              options = list(scrollX = TRUE,
                             order = list(list(1, 'asc')),
                             autoWidth = TRUE))
```

View locations on a map. 

```{r}
leaflet::leaflet()  |> 
  leaflet::addTiles() |> 
  leaflet::addMarkers(lng = deltawqAQ::aq_all_locations$longitude, 
             lat = deltawqAQ::aq_all_locations$latitude, 
             label = deltawqAQ::aq_all_locations$cdec_code)
```

#### Look at available parameters
Once you have picked your location, take a look at the parameters you would like to query. The `aq_get_location_parameters` function will connect to the Aquarius database based on your credentials, then retrieve all parameters associated with the location you have selected. 

We have selected the `SJW` location for this example. You will need `parameter_name` for your time series query, so identify the correct value for the next step.

```{r}
param_list <- aq_get_location_parameters(cdec_code = "SJW")
DT::datatable(param_list, 
              rownames = FALSE,
              options = list(scrollX = TRUE, autoWidth = TRUE))
```

### Start with parameter

Instead of starting with location, you may also start by figuring out the parameter code(s) of interest. 

```{r}
DT::datatable(deltawqAQ::aq_all_parameters, 
              rownames= FALSE,
              options = list(scrollX = TRUE,
                             order = list(list(1, 'asc'))))
```

#### Look at available locations

From there you can look at which locations have data for your parameter of interest

```{r}
wt_locations <- aq_get_parameter_locations("Water Temp")
DT::datatable(wt_locations, 
              rownames= FALSE,
              options = list(scrollX = TRUE,
                             autoWidth = TRUE))
```

### Query data

Now you can run your query for your location and parameter of interest. You can enter a datetime or a date for the `query_from` and `query_to` fields. The function defaults to include data from 2 weeks prior to the current date. If you don't enter query dates, the function defaults to include data from 2 weeks prior to the current date.

```{r}
ts <- aq_get_ts(cdec_code = "SJW", 
                parameter = "Water Temp", 
                query_from = "2025-12-01 00:00:00", 
                query_to = "2026-01-01 00:00:00")
```

```{r}
DT::datatable(ts, 
              rownames = FALSE, 
              options = list(scrollX = TRUE, autoWidth = TRUE))
```

### Summarize data
You may want to condense the data by creating a dataset with daily averages. 

```{r, warning = FALSE, message = FALSE}
ts_daily <- ts |> 
  mutate(date = lubridate::date(datetime)) |>
  group_by(date, cdec_code) |>
  summarize(wt_daily = round(mean(value),2)) |>
  ungroup()
```

```{r}
DT::datatable(ts_daily, 
              rownames = FALSE,
              options = list(scrollX = TRUE))
```

### Visualize data
Now you can make a simple plot for daily average values for the time series.

```{r, fig.height = 4, fig.width = 6, fig.title = ts_daily$aq_location_name[1], fig.alt = "Plot of daily average water temperature at SJW"}

ggplot(ts_daily) + 
  geom_point(aes(x = date, y = wt_daily)) +
  geom_line(aes(x = date, y = wt_daily)) +
  labs(x = "Date", y = "Water Temp (°C)") +
  theme_bw()+
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 11))
```

### Write data

You can then write data to your local drive. Specify the filepath of interest and customize the filename if desired. 

```{r, eval = FALSE}
write.csv(ts_daily, 
          here::here(paste0("ts_SJW_wt.csv")))
```

## Multiple parameters for one location

Now we can try an example for getting multi-parameter data. 

### Explore locations

As in the first example, you can start by figuring out the location code(s) of interest. Functions will take `cdec_code`, `location_id`, or `aquarius_id`, with most users tending to be most familiar with `cdec_code`. 

Use the `aq_all_locations` data object to view all the available locations. This object gets generated in the package, and locations should be stable once all locations become telemetered.

```{r}
DT::datatable(deltawqAQ::aq_all_locations, 
              rownames = FALSE,
              options = list(scrollX = TRUE,
                             order = list(list(0, 'asc')),
                             autoWidth = TRUE))
```

Let's select SJW. 

```{r}
location = "SJW"
```

Once you have picked your location, take a look at the parameters you would like to query. The `aq_get_location_parameters` function will connect to the Aquarius database based on your credentials, then retrieve all parameters associated with the location you have selected. You will need `parameter_name` for your time series query, so identify the correct value for the next step.

### Select parameter names
Now use `aq_get_location_parameters` to figure out the parameter names for your location.

```{r}
DT::datatable(aq_get_location_parameters(cdec_code = location), 
              rownames = FALSE,
              options = list(scrollX = TRUE, 
                             autoWidth = TRUE))
```

### Query data 
Now that you have your location and parameters you can use the `aq_get_ts` function to run your query for your the timeframe of interest. You can enter a datetime or a date for the `query_from` and `query_to` fields. If you don't enter query dates, the function defaults to include data from 2 weeks prior to the current date.

#### Select parameters and dates
```{r}
params = c("Turbidity, Form Neph", "Water Temp", "Sp Cond", "CHL RFU", "pH")
start_date = lubridate::today()-21
end_date = lubridate::today()
```

#### Query and look at data
```{r, message = FALSE, warning = FALSE}
multi_param_ts <- aq_get_ts(
  cdec_code = location, 
  parameter = params, 
  query_from = start_date, 
  query_to = end_date
  )

print(paste("Parameters with data:", paste(unique(multi_param_ts$parameter_name), collapse = ", ")))
```

```{r}
DT::datatable(multi_param_ts, 
              rownames = FALSE,
              options = list(scrollX = TRUE, 
                             autoWidth = TRUE))
```


### Summarize by daily values
You may want to condense the data by creating a dataset with daily averages. 

```{r}
multi_param_daily <- multi_param_ts  |> 
  mutate(date = lubridate::date(datetime)) |> 
  group_by(date, cdec_code, location_id, parameter_name) |> 
  summarize(daily_avg = round(mean(value),2)) |> 
  ungroup()
```

```{r}
DT::datatable(multi_param_daily, 
              rownames = FALSE,
              options = list(scrollX = TRUE))
```


### Visualize time series for each parameter

If you'd like to visualize your data, here are some examples of plots you can make with your data. 

#### Daily Averages
```{r fig.height = 7, fig.width = 6, fig.alt = "Plot of daily average values across different parameters at San Joaquin West (SJW)"}
ggplot(multi_param_daily) +
  geom_point(aes(x=date, y=daily_avg, color = parameter_name)) +
  geom_line(aes(x=date, y=daily_avg, color = parameter_name)) +
  facet_wrap(~parameter_name, nrow = nrow(multi_param_daily), scales = "free_y") +
  labs(x = "Date", y = "Daily Average Value", title = paste("Daily Average Water Quality for", location))+
  theme_bw() +
  theme(strip.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 11))
```

#### 15 minute data
```{r, fig.height = 7, fig.width = 6, message = FALSE, warning = FALSE, fig.alt = "Plot of 15-minute values across different parameters at San Joaquin West (SJW)"}
ggplot(multi_param_ts) +
  geom_line(aes(datetime, value, color = parameter_name)) +
  facet_wrap(~parameter_name, nrow = length(unique(multi_param_ts$parameter_name)), scales = "free_y") +
  labs(x = "Datetime", y = "Value", title = paste("15-minute Water Quality for", location))+
  theme_bw() +
  theme(strip.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 11))
```

### Write data
You can then write data to your local drive. Specify the filepath of interest and customize the filename if desired. This example writes the averaged data. 

```{r, eval=FALSE}
param_label <- paste(unique(multi_param_ts$parameter_name), collapse = "-")

write.csv(multi_param_daily, 
          here::here(paste0("ts_", location, "_",param_label,"_",  start_date, "-", end_date)))
```

## Multiple locations for one parameter

### Explore parameter names

Another option is to start by figuring out the parameter(s) of interest. Functions will use `parameter_name` for queries.

Use the `aq_all_parameters` data object to view all the available parameters. This object gets generated in the package, and parameters should be stable once all locations become telemetered.

```{r}
DT::datatable(deltawqAQ::aq_all_parameters, 
              rownames = FALSE,
              options = list(scrollX = TRUE,
                             order = list(list(1, 'asc'))))
```

### Select locations 

Now use `aq_get_parameter_locations` function to figure out the locations that contain data for your parameter.

```{r}
turbidity_locations <- aq_get_parameter_locations("Turbidity, Form Neph")
DT::datatable(turbidity_locations, 
              rownames = FALSE,
              options = list(scrollX = TRUE, autoWidth = TRUE))
```

Now select your locations from the list. 

```{r}
locations <- c("SJW", "SGG", "RYF", "TOE")
```

### Query data 
Now that you have your locations and parameter you can use the `aq_get_ts` function to run your query for the timeframe of interest. You can enter a datetime or a date for the `query_from` and `query_to` fields. If you don't enter query dates, the function defaults to include data from 2 weeks prior to the current date.

#### Select parameter and dates
```{r}
param = "Turbidity, Form Neph"
startdate = "2025-12-01"
enddate = lubridate::today()
```

#### Query and look at data
```{r, message = FALSE, warning = FALSE}
multi_location_ts <- aq_get_ts(cdec_code = locations, 
                                         parameter = param, 
                                         query_from = startdate, 
                                         query_to = enddate)

print(paste("locations with data:", paste(unique(multi_location_ts$cdec_code), collapse = ", ")))
```

```{r}
DT::datatable(multi_location_ts, 
              rownames = FALSE,
              options = list(scrollX = TRUE, autoWidth = TRUE))
```

### Summarize by daily values
You may want to condense the data by creating a dataset with daily averages. 

```{r}
multi_location_daily <- multi_location_ts |> 
  mutate(date = lubridate::date(datetime)) |> 
  group_by(date, cdec_code, location_id, parameter_name) |> 
  summarize(daily_avg = round(mean(value),2)) |> 
  ungroup()
```

```{r}
DT::datatable(multi_location_daily, 
              rownames = FALSE,
              options = list(scrollX = TRUE))
```


### Visualize time series for each location

If you'd like to visualize your data, here are some examples of plots you can make with your data. 

#### Daily Averages

```{r, fig.height = 7, fig.width = 6, fig.alt = "Plot of daily average turbidity across stations"}
ggplot(multi_location_daily) +
  geom_point(aes(x=date, y=daily_avg, color = cdec_code)) +
  geom_line(aes(x=date, y=daily_avg, color = cdec_code)) +
  facet_wrap(~cdec_code, nrow = nrow(multi_location_daily), scales = "free_y") +
  labs(x = "Date", y = "Daily Average Value", title = paste("Daily Average", param, "across stations"))+
  theme_bw() +
  theme(strip.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 11))
```

#### 15 minute data

Include a threshold line if useful.

```{r, message = FALSE, warning = FALSE, fig.height = 7, fig.width = 6, fig.alt = "Plot of 15-minute turbidity data across multiple stations with a threshold at 12 FNU."}
ggplot(multi_location_ts) +
  geom_line(aes(datetime, value, color = cdec_code)) +
  geom_hline(yintercept = 12, linetype = "dashed") + 
  facet_wrap(~cdec_code, nrow = nrow(multi_location_ts), scales = "free_y") +
  labs(x = "Datetime", y = param,  title = paste("15-minute", param, "across stations"))+
  theme_bw() +
  theme(strip.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.position = "none",
        axis.text = element_text(size = 11))
```

### Write data

You can then write data to your local drive. Specify the filepath of interest and customize the filename if desired. 

```{r, eval = FALSE}
loc_label <- paste(unique(multi_location_ts$cdec_code), collapse = "-")
write.csv(multi_location_daily, 
          here::here(paste0("ts_", param, "_",sta_label,"_",  start_date, "-", end_date)))
```

